---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(nhanesA)
library(tidyverse)
```


# STUDY 1 IDEAS

A:

B:Height vs. wanting to lose weight. 
BMX_J and WHQ_J

C:Height vs. Current Health status
BMX_J and HSQ_J

D:Smoking vs. Military History
SMQ_J(SMQ020) and DEMO_J(DMQMILIZ) 

E:Have little interest in doing things vs. How healthy is diet

# STUDY 2 IDEAS

Outcome of interest: Weight(BMX_J)

Predictors: Eating Habits(DBQ_J(DBQ700))

            My Plate Knowledge(DBQ_J(CBQ596))
            
            Meals Not Home Prepared(DBQ_J(DBD895))
            
            Money Spent at supermarket(CBQ_J(CBD071))
            
            Annual Household Income(DEMO_J(INDHHIN2))





```{r}
demo_raw <- nhanes('DEMO_J')
bmx_raw <- nhanes('BMX_J')
whq_raw <- nhanes('WHQ_J')
hsq_raw <- nhanes('HSQ_J')
smq_raw <- nhanes('SMQ_J')
dpq_raw <- nhanes('DPQ_J')
dbq_raw <- nhanes('DBQ_J')
cbq_raw <- nhanes('CBQ_J')
ocq_raw <- nhanes('OCQ_J')
```

```{r}
temp1 <- left_join(demo_raw, bmx_raw, by = "SEQN")

temp2 <- left_join(temp1, whq_raw, by = "SEQN")

temp3 <- left_join(temp2, hsq_raw, by = "SEQN")

temp4 <- left_join(temp3, smq_raw, by = "SEQN")

temp5 <- left_join(temp4, dpq_raw, by = "SEQN")

temp6 <- left_join(temp5, dbq_raw, by = "SEQN")
dim(temp6)
```

```{r}
saveRDS(temp6, file = "NHANES_study1.Rds")
```


```{r}
temp1 <- left_join(demo_raw, dbq_raw, by = "SEQN")
temp2 <- left_join(temp1, cbq_raw, by = 'SEQN')
temp3 <- left_join(temp2, bmx_raw, by = "SEQN")
temp4 <- left_join(temp3, ocq_raw, by = "SEQN")



```

```{r}
dim(study2)
```

```{r}
saveRDS(temp4, file = "NHANES_study2.Rds")
```


Notes from presentation:
Study 1: Analysis B

  Include units for measurements for description
  
  Remove legend from boxplot 

Study2 : Make sure to use glance to describe model and assess r squared/nobs

Study2 : Look at headings for study 2 from example and re header the study 2



